---
title: "Math 204 - Machine Problem 1"
author: "Francis Dave Cabanting"
date: "10/4/2020"
output: 
  html_document:
    keep_md: true
    toc: true
---

## Part I
#### 1.
Create a variable in R that contains the total number of Active Case and assign it to `total`.
```{r}
asym <- 326
mild <- 67
seve <- 7
crit <- 1

total <- asym + mild + seve + crit
total
```
#### 2.
Create a numeric vector in R that contains the distribution of COVID-19 cases `{326, 67, 7, 1}` and assign them to `dist_Cases`. Assign their status category `{asymptomatic, mild, severe and critical}` as names of the vector elements.
```{r}
dist_Cases <- c('asymptomatic' = asym,
                'mild' = mild,
                'severe' = seve,
                'critical' = crit)
dist_Cases
```
#### 3. 
Using the created numeric vector in part I.2, select the values representing the mild and asymptomatic cases and assign them to `mild_2_asymtomatic`.

```{r}
mild_2_asymptomatic <- dist_Cases[c('mild','asymptomatic')]
mild_2_asymptomatic
```
#### 4.
In the data province with new cases, we only have Davao City and Davao del Norte. Davao del Sur, Davao Oriental, Davao de Oro and Davao Occidental have no new COVID-19 cases. Create an object in R (numeric vector) and assign it to `newcases`. Include those provinces with no new cases.

```{r}
provinces <- c('Davao City',
               'Davao del Norte',
               'Davao del Sur',
               'Davao Oriental',
               'Davao de Oro',
               'Davao Occidental')
newcases <- c(7,7,0,0,0,0)
names(newcases) <- provinces
newcases
```
#### 5.
Print the object `newcases` by selecting only the provinces with new COVID-19 cases.
```{r}
newcases[newcases>0]
```
## Part II

#### a.
Create an R object or matrix named `breakdown` that will hold the COVID-19 total active cases, total deaths, and total recoveries for the different provinces and city in Davao Region. Assign appropriate column names `{Total Active, Total Deaths, Total Recoveries}` and row names `{names of provinces and city}` to the matrix.

```{r}
provs <- c('Davao City',
           'Davao de Oro',
           'Davao del Norte',
           'Davao del Sur',
           'Davao Occidental',
           'Davao Oriental')
categs <- c('Total Active Cases',
           'Total Deaths',
           'Total Recoveries')
breakdown <- matrix(c(214,57,1200,
                      66,1,117,
                      68,4,203,
                      31,4,101,
                      4,2,36,
                      18,1,196),
                    nrow = 6,
                    ncol = 3,
                    byrow = TRUE)
colnames(breakdown) <- categs
rownames(breakdown) <- provs

breakdown
```


#### b.
Get a subset of the matrix `breakdown` by selecting only the data for Davao City.
```{r}
breakdown['Davao City',]
```
#### c.
Get a subset of the matrix by selecting those whose number of deaths is below `4`.
```{r}
breakdown[breakdown[,'Total Deaths']<4,]
```
#### d.
For the whole Davao Region get the total number of active cases, total deaths and total recoveries as of September 8, 2020, and assign it to the object `sumDavao`.
```{r}
sumDavao <- colSums(breakdown)

sumDavao
```
#### e.
Create a data frame using the same data presented in Table 1 and 2 above and assign it to `covid_data`. Note: Some information in Table 1 do not have provincial/city breakdowns, hence, you cannot include them in the data frame.

```{r}
covid_data <- data.frame(breakdown,
                         'New Cases' = newcases)
covid_data
```

#### f.
Insert a column to the data frame `covid_data` by adding a column for new cases `[TRUE/FALSE]`.
```{r}
covid_data[['Has New Cases']] <- covid_data['New.Cases']>0
covid_data
```
## Part III

You are given a sample of 50 COVID-19 cases based on the records release by DOH. The data in excel format is available in the UVE as an attachment to machine problem 1.



#### i. 
Import the dataset to R.

#### ii. 
Convert the data into a tibble data frame and assign it to `covid_raw`.

#### iii. 
Create a new data frame `above_60` by selecting the patients who are above 60 years old.

#### iv. 
Extract the patients who is above 60 years old who already recovered from COVID-19. Assign the new data frame to `recovery_60`.

#### v. 
Create a new data frame `no_signs` that contains the asymtomatic patients or patients with `NA` values for the listed symptoms for COVID-19.


